---
title: "DeKalb County by Zipcode Methodology"
output:
  github_document: default
---
July 9, 2020

```{r setup}
#Load required libraries

library(tidyverse)
library(modelr)
library(ggrepel)

```

The [DeKalb County Board of Health](https://www.dekalbhealth.net/covid-19dekalb/) has released data on local COVID-19 cases, sorted by ZIP Code. This information has not been shared in a format that is suitable for data analysis, so I copied-and-pasted the raw numbers into a text file, which I then imported and tidied. 

```{r}
#Import and tidy data from DeKalb County Board of Health 
#https://www.dekalbhealth.net/covid-19dekalb/

dekalb <- read_lines("Dekalb.txt", skip_empty_rows = TRUE)
dekalb <- dekalb[dekalb != "\t"]
dekalb <- dekalb[6:(length(dekalb))]
dekalb <- matrix(dekalb, ncol = 5, byrow = TRUE) %>%
  as_tibble(.name_repair = "unique") %>%
  select(...1, ...3) %>%
  rename(ZIP = ...1,
         current_count = ...3,)
```

The U.S. Census Bureau's most recent American Community Survey 5-year estimate, released in 2018, provides estimates at the ZIP Code level (or rather ZCTA) for a variety of statistics on the U.S. population. The 2010 Census provides population counts for each ZCTA, information we need to calculate cases per capita.

```{r}
#Import data from US Census American Community Survey 2018 5-year estimate

zcta <-
  read_csv("ACSST5Y2018.S1901_data_with_overlays_2020-07-04T190723.csv",
           skip = 1, na = "null") %>%
  mutate(ZIP = str_extract(`Geographic Area Name`,
                           ".....$")) %>%
  select(ZIP, `Estimate!!Households!!Median income (dollars)`)

#Import population data from 2010 Census. I originally used data for the entire 
#ZCTA, but this doesn't work because some of these ZCTAs do not fall entirely 
#within DeKalb. We need to figure out how many people live in DeKalb, not in the 
#ZCTA as a whole. Fortunately the Census Bureau provides a file that specifies 
#the relation between ZCTAs and counties, and this document gives us the numbers 
#we need.  

zcta_pop <- read_csv("zcta_county_rel_10.txt") %>%
  filter(STATE == "13" & COUNTY == "089") %>%
  select(ZIP = ZCTA5, total_population = POPPT)
```

Now that we have the data from the county Board of Health and the Census Bureau, we can combine them into a single data frame and calculate cases per thousand population. 

```{r}
#Join ACS data with local COVID-19 data and specify column types

dekalb <- dekalb %>%
  inner_join(zcta) %>%
  mutate(
    current_count = as.numeric(current_count),
    median_income = as.numeric(`Estimate!!Households!!Median income (dollars)`)
  ) %>%
  select(!(`Estimate!!Households!!Median income (dollars)`))

#Add population data to Dekalb tibble and calculate per capita data

dekalb <- dekalb %>%
  inner_join(zcta_pop) %>%
  mutate(cases_per_thousand = current_count / (total_population / 1000))

#clear redundant tibbles from memory

remove(zcta)
remove(zcta_pop)
```

Now that these data have been tidied, combined, and calculated, we can take a look and see if there's a pattern.

```{r}
#Now let's make a plot to see if there is a correlation between income and disease

ggplot(data = dekalb,
       mapping = aes(median_income, cases_per_thousand)) +
  geom_point() +
  geom_smooth()
```

Before I corrected the data to account for ZCTAs not falling entirely within the county, it seemed like there was a correlation between median income and COVID-19, but I was wrong! But let's go ahead and fit the data to a linear model anyway (since it's already done). 

```{r}
#It looks like there's a correlation. Let's try modeling it

mod_covid <- lm(cases_per_thousand ~ median_income, dekalb)
summary(mod_covid)
```

That confirms there's nothing there!
