---
title: "Metro Atlanta COVID"
output: 
  html_document: 
    keep_md: yes
---

```{r setup, include = FALSE}

#Load the needed packages

library(tidyverse)

#Grab the raw data from the NY Times

counties <- read_csv(
        "https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv", 
                     col_types = cols(
        date = col_date(format = ""),
        county = col_character(),
        state = col_character(),
        fips = col_character(),
        cases = col_double(),
        deaths = col_double()
))

#Calculate the total number of cases and deaths in the US

usa <- counties %>%
        group_by(date) %>%
        summarise("all_cases" = sum(cases),
                  "all_deaths" = sum(deaths)) %>%
        mutate("new_deaths" = all_deaths - lag(all_deaths, default = 0),
               "new_cases" = all_cases - lag(all_cases, default = 0))

#Calculate the total number of cases and deaths in Georgia

georgia <- counties %>%
        filter(state == "Georgia") %>%
        group_by(date) %>%
        summarise("all_cases" = sum(cases),
                  "all_deaths" = sum(deaths)) %>%
        mutate("new_deaths" = all_deaths - lag(all_deaths, default = 0),
               "new_cases" = all_cases - lag(all_cases, default = 0))

#Filter down to the five core metro Atlanta counties

metro_atl <- counties %>%
        filter(state == "Georgia" & (county == "DeKalb" | county == "Gwinnett" | 
                county == "Fulton" | county == "Cobb" | county == "Clayton"))

#Add the census counts for the metro Atlanta counties

atl_counties_population <- enframe(c("DeKalb" = 756558, "Gwinnett" = 927781, 
        "Fulton" = 1050114, "Cobb" = 756865, "Clayton" = 289615), 
        name = "county", value = "county_population")
```

I wanted a better visualization of local Covid-19 data than I've been able to 
find, particularly in terms of the distribution of cases and deaths
as they change over time. The charts below track confirmed cases and deaths in 
the core Metro Atlanta counties, both in absolute and relative terms. The 
subsequent charts track new cases and deaths in Metro Atlanta, Georgia, and the 
United States.

All data come from the New York Times' ongoing [repository](https://github.com/nytimes/covid-19-data) 
of Covid-19 cases and deaths in the United States.

It's hard to know how to interpret these numbers since there are major known
unknowns. Neither confirmed cases nor deaths can be said to be reliable counts
of the true numbers.

# Metro Atlanta

## Total Cases by County

```{r atl_cases_absolute, echo = FALSE}

#Visualize Metro Atlanta county case data in absolute terms

        ggplot(data = metro_atl) +
        geom_point(mapping = aes(x = date, y = cases, color = county)) +
        geom_line(mapping = aes(x = date, y = cases, color = county)) 
```

## Total Deaths by County

```{r atl_deaths_absolute, echo = FALSE}

#Visualize Metro Atlanta county case data in absolute terms

        ggplot(data = metro_atl, mapping = aes(x = date, y = deaths, 
                color = county)) +
        geom_point() +
        geom_line()
```

## Cases per Thousand by County

```{r atl_relative_cases, echo = FALSE}

#Visualize Metro Atlanta County cases in relative terms

metro_atl %>%
        mutate(cases_per_thousand = (cases / 
                atl_counties_population$county_population) * 1000) %>%
ggplot(mapping = aes(x = date, y = cases_per_thousand, color = county)) +
        geom_point() +
        geom_line()
```

## Deaths per Thousand by County

```{r alt_relative_deaths, echo = FALSE}

#Visualize Metro Atlanta county deaths in relative terms

metro_atl %>%
        mutate(deaths_per_thousand = (deaths / 
                atl_counties_population$county_population) * 1000) %>%
ggplot(mapping = aes(x = date, y = deaths_per_thousand, color = county)) +
        geom_point() +
        geom_line()

```

## New Cases in Metro Atlanta

```{r atl_new_cases, echo = FALSE}

#Visualize new cases for entire metro area

        metro_atl %>%
                group_by(date) %>%
                summarise("all_cases" = sum(cases)) %>%
                mutate("new_cases" = 
                               all_cases - lag(all_cases, default = 0)) %>%
                ggplot(mapping = aes(x = date, y = new_cases)) +
                geom_col() +
                geom_smooth(se = FALSE)
```

## New Deaths in Metro Atlanta

```{r atl_new_deaths, echo = FALSE}

#Visualize new deaths for entire metro area

        metro_atl %>%
                group_by(date) %>%
                summarise("all_deaths" = sum(deaths)) %>%
                mutate("new_deaths" = 
                               all_deaths - lag(all_deaths, default = 0)) %>%
                ggplot(mapping = aes(x = date, y = new_deaths)) +
                geom_col() +
                geom_smooth(se = FALSE)
```


# Georgia

## New Cases

```{r georgia_new_cases, echo = FALSE}
ggplot(georgia, mapping = aes(x = date, y = new_cases), lab) +
        geom_col() +
        geom_smooth(se = FALSE, span = .5)
        
```

## New Deaths

```{r georgia_new_deaths, echo = FALSE}
ggplot(georgia, mapping = aes(x = date, y = new_deaths)) +
        geom_col() +
        geom_smooth(se = FALSE, span = .5)
```


# United States

## New Cases

```{r us_cases, echo = FALSE}
ggplot(usa, mapping = aes(x = date, y = new_cases)) +
        geom_col() +
        geom_smooth(se = FALSE, span = .25)
```

## New Deaths

```{r us_deaths, echo = FALSE}
ggplot(usa, mapping = aes(x = date, y = new_deaths)) +
        geom_col() +
        geom_smooth(se = FALSE, span = .25)

```

