---
title: "Metro Atlanta COVID"
output:
  github_document: default
---

```{r setup, include = FALSE}

#Load the needed packages


library(tidyverse)
library(zoo)

knitr::opts_chunk$set(
  message = FALSE,
  echo = FALSE,
  warning = FALSE,
  out.width = "70%",
  fig.align = "center"
)

#Grab the raw data from the NY Times

counties <- read_csv(
  "https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv",
  col_types = cols(
    date = col_date(format = ""),
    county = col_character(),
    state = col_character(),
    fips = col_character(),
    cases = col_double(),
    deaths = col_double()
  )
)

#Calculate the total number of cases and deaths in the US

usa <- counties %>%
  group_by(date) %>%
  summarise("all_cases" = sum(cases),
            "all_deaths" = sum(deaths)) %>%
  mutate(
    "new_deaths" = all_deaths - lag(all_deaths, default = 0),
    "new_cases" = all_cases - lag(all_cases, default = 0)
  ) %>%
  mutate("weekly_mean_deaths" = rollmean(
    new_deaths,
    7,
    na.rm = TRUE,
    fill = 0,
    align = "right"
  )) %>%
  mutate("weekly_mean_cases" = rollmean(
    new_cases,
    7,
    na.rm = TRUE,
    fill = 0,
    align = "right"
  ))

#Calculate the total number of cases and deaths in Georgia

georgia <- counties %>%
  filter(state == "Georgia") %>%
  group_by(date) %>%
  summarise("all_cases" = sum(cases),
            "all_deaths" = sum(deaths)) %>%
  mutate(
    "new_deaths" = all_deaths - lag(all_deaths, default = 0),
    "new_cases" = all_cases - lag(all_cases, default = 0)
  ) %>%
  mutate("weekly_mean_deaths" = rollmean(
    new_deaths,
    7,
    na.rm = TRUE,
    fill = 0,
    align = "right"
  )) %>%
  mutate("weekly_mean_cases" = rollmean(
    new_cases,
    7,
    na.rm = TRUE,
    fill = 0,
    align = "right"
  ))

#Filter down to the five core metro Atlanta counties

metro_atl <- counties %>%
  filter(
    state == "Georgia" & (
      county == "DeKalb" | county == "Gwinnett" |
        county == "Fulton" |
        county == "Cobb" | county == "Clayton"
    )
  )

#Add the census counts for the metro Atlanta counties

atl_counties_population <-
  enframe(
    c(
      "DeKalb" = 756558,
      "Gwinnett" = 927781,
      "Fulton" = 1050114,
      "Cobb" = 756865,
      "Clayton" = 289615
    ),
    name = "county",
    value = "county_population"
  )

#Define functions to draw plots

case_chart <- function(region, title) {
  ggplot(region, mapping = aes(x = date)) +
    geom_col(mapping = aes(y = new_cases, fill = "new daily cases")) +
    geom_line(mapping = aes(y = weekly_mean_cases,
                            color = "7-day average")) +
    labs(
      title = title,
      caption = "Data from The New York Times",
      x = "Date",
      y = "New Cases"
    ) +
    scale_color_manual(values = c("7-day average" = "#800014")) +
    scale_fill_manual(values = c("new daily cases" = "gray40")) +
    theme(legend.title = element_blank(),
          legend.position = "top") +
    scale_x_date(limits = as.Date(c("2020-03-01", Sys.time())))
}

death_chart <- function(region, title) {
  ggplot(region, mapping = aes(x = date)) +
    geom_col(mapping = aes(y = new_deaths,
                           fill = "new daily deaths")) +
    geom_line(mapping = aes(y = weekly_mean_deaths,
                            color = "7-day average")) +
    labs(
      title = title,
      caption = "Data from The New York Times",
      x = "Date",
      y = "New Deaths"
    ) +
    scale_color_manual(values = c("7-day average" = "black")) +
    scale_fill_manual(values = c("new daily deaths" = "#800014")) +
    theme(legend.title = element_blank(),
          legend.position = "top") +
    scale_x_date(limits = as.Date(c("2020-03-01", Sys.time())))
}
```

The charts below track confirmed cases and deaths in 
the core Metro Atlanta counties, both in absolute and relative terms. The 
subsequent charts track new cases and deaths in Metro Atlanta, Georgia, and the 
United States.

All data come from the New York Times' ongoing [repository](https://github.com/nytimes/covid-19-data) 
of COVID-19 cases and deaths in the United States.

It's hard to know how to interpret these numbers since there are major known
unknowns. Neither confirmed cases nor deaths can be said to be reliable counts
of the true numbers.

Data current as of `r format(Sys.time(), '%B %d, %Y')`.

# Metro Atlanta

```{r atl_cases_absolute}
#Visualize Metro Atlanta county case data in absolute terms

ggplot(data = metro_atl) +
  geom_line(mapping = aes(x = date, y = cases, color = county)) +
  labs(
    title = "Total Cases by County",
    caption = "Data from The New York Times",
    color = "County",
    x = "Date",
    y = "Total Cases"
  ) +
  theme(legend.position = c(.15, .7)) +
  scale_x_date(limits = as.Date(c("2020-03-01", Sys.time())))
```


```{r atl_deaths_absolute}
#Visualize Metro Atlanta county death data in absolute terms

ggplot(data = metro_atl,
       mapping = aes(x = date, y = deaths,
                     color = county)) +
  geom_line() +
  labs(
    title = "Total Deaths by County",
    caption = "Data from The New York Times",
    color = "County",
    x = "Date",
    y = "Total Deaths"
  ) +
  theme(legend.position = c(.15, .7)) +
  scale_x_date(limits = as.Date(c("2020-03-01", Sys.time())))
```


```{r atl_relative_cases}

#Visualize Metro Atlanta County cases in relative terms

metro_atl %>%
  mutate(cases_per_thousand = (cases /
                                 atl_counties_population$county_population) * 1000) %>%
  ggplot(mapping = aes(x = date, y = cases_per_thousand, color = county)) +
  geom_line() +
  labs(
    title = "Cases per Thousand by County",
    caption = "Data from The New York Times",
    color = "County",
    x = "Date",
    y = "Cases per Thousand"
  ) +
  theme(legend.position = c(.15, .7)) +
  scale_x_date(limits = as.Date(c("2020-03-01", Sys.time())))
```

```{r alt_relative_deaths}

#Visualize Metro Atlanta county deaths in relative terms

metro_atl %>%
  mutate(deaths_per_thousand = (deaths /
                                  atl_counties_population$county_population) * 1000) %>%
  ggplot(mapping = aes(x = date, y = deaths_per_thousand, color = county)) +
  geom_line() +
  labs(
    title = "Deaths per Thousand by County",
    caption = "Data from The New York Times",
    color = "County",
    x = "Date",
    y = "Deaths per Thousand"
  ) +
  theme(legend.position = c(.15, .7)) +
  scale_x_date(limits = as.Date(c("2020-03-01", Sys.time())))

```

```{r atl_new_cases}

#Visualize new cases for entire metro area

metro_atl %>%
  group_by(date) %>%
  summarise("all_cases" = sum(cases)) %>%
  mutate("new_cases" =
           all_cases - lag(all_cases, default = 0)) %>%
  mutate("weekly_mean_cases" = rollmean(
    new_cases,
    7,
    na.rm = TRUE,
    fill = 0,
    align = "right"
  )) %>%
  case_chart(., "New Cases in Metro Atlanta")
```

```{r atl_new_deaths}

#Visualize new deaths for entire metro area

metro_atl %>%
  group_by(date) %>%
  summarise("all_deaths" = sum(deaths)) %>%
  mutate("new_deaths" =
           all_deaths - lag(all_deaths, default = 0)) %>%
  mutate("weekly_mean_deaths" = rollmean(
    new_deaths,
    7,
    na.rm = TRUE,
    fill = 0,
    align = "right"
  )) %>%
  death_chart(., "New Deaths in Metro Atlanta")
```


# Georgia

```{r georgia_new_cases}
case_chart(georgia, "New Cases in Georgia")
```

```{r georgia_new_deaths}
death_chart(georgia, "New Deaths in Georgia")
```


# United States

```{r us_cases}
case_chart(usa, "New Cases in the United States")
```

```{r us_deaths}
death_chart(usa, "New Deaths in the United States")
```

